---
title: "01-exercises"
author: "Saqib Ali"
date: "April 2, 2017"
output: html_document
---


## Exercise 1:

Write a function (f) that takes a vector of numbers, **x**, and returns a vector of numbers such that each element containing the product of every element of x except the element of x with the same index. 


### Example 

    > x <- c( 1, 5, 2, 8 ) 
    > f(x) 
    [1] 80 16 40 10
    # 5*2*8, 1*2*8, 1*5*8, 1*2*5  


### Solution 

```{r}
x <- c( 1, 5, 2, 8 ) 
f <- function(x) {
    prod(x)/x 
}
f(x)

```
    


## Exercise 2

Write a function f(x) to accept an integer vector, and returns a vector with those numbers except for: 
   multiples of 3 replaced by "Fizz"? 
   multiples of 5 replaced by "Buzz" 
   multiples of 3 and 5 replaced by "FizzBuzz""
   
### Example

    > x <- 1:20
    > f(1:20) 
    # 1 2 Fizz 4 Buzz Fizz 7 8 Fizz Buzz 11 Fizz 13 14 FizzBuzz 16 17 Fizz 19 Buzz


### Solution

```{r}

x <- 1:20
######## Method 1
f1 <- function(x) {
  x[as.numeric(x)%%3==0 & as.numeric(x)%%5==0] <- "FizzBuzz"
  x[as.numeric(x)%%3==0] <- "Fizz"
  x[as.numeric(x)%%5==0] <- "Buzz"
  x  
}

###### Method 2
f2 <- function(x){
  y <- c()
  for (i in 1:length(x)){
    if ( (i%%3==0) & (i%%5==0) ) {y <- append(y,"FizzBuzz") }
    else if (i%%3==0) {y <- append(y,"Fizz") }
    else if (i%%5==0) {y <- append(y,"Buzz") }
    else {y <- append(y,i) }
  }
  y
}


#### Testing
f1(x)
f2(x)


all.equal(f1(x), f2(x))
```
