---
title: 'NYCFlights: Arrival Delay Logictic Model'
author: "Saqib Ali"
date: "May 1st, 2017"
output:
  html_document: default
  pdf_document: default
---




## Logsitic and Inverse Logistic Transformation 

- Write an R function for the logistic function. The function should accept a `numeric` vector with values `[-Inf,Inf]` and produce a numeric vector in the the range `[0,1]`.

- Plot the logistic function from  `[-10,10]`

- Write a R function for the inverse logistic function. The function should accept a `numeric` vector with values `[0,1]` and prodcuce a numeric vector in the range `[-Inf,Inf]`

- Plot the Inverse Logistic function from `[0,1]`


**Hint:** For plotting curves see `?graphics::curve` or `?ggplot2::stat_function`

```{r "Logistic and Inverse Logistic" }

logistic <- function(x) { 
  1/(1+exp(-x))
}

inv_logistic <- function(x) { 
  -log((1-x)/x)
}

x <- -10:10
curve(logistic, -10, 10)
curve(inv_logistic, 0, 1)


```




# NYCFlights Model

Using the rectangular data that you created from the earlier assignment and following theexample from the text and class, create a model for arr_delay >= 22 minutes. Describe/Explain each of the steps and show all work.

KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file to your repository.   


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library("tidyverse")
library("ggplot2")
library("lubridate")
library("dplyr")
library("data.table")

library(readr)
library(ggplot2)
library(magrittr)
library(lubridate)
library(stringr)
library(dplyr)
#library(caret)
library(InformationValue)


library(data.table)


flightsDT  <- fread("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/flights.csv") 
planesDT   <- fread("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/planes.csv")  %>% setkey(tailnum)
airportsDT <- fread("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/airports.csv")  %>% setkey(faa) 
weatherDT  <- fread("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/weather.csv") %>%   setkey(origin,year,month,day,hour)



```



## Join the datasets and analyze the structure
```{r}
YX <- flightsDT 
YX %<>% merge( planesDT, all.x = TRUE, by='tailnum', suffixes=c('','.pl') )
YX %<>% merge( weatherDT, all.x = TRUE, by=c('origin','year','month','day','hour'), suffixes=c('','.we') )
YX %<>% merge( airportsDT, all.x = TRUE, by.x='origin', by.y='faa', suffixes=c('','.orig') )
YX %<>% merge( airportsDT, all.x = TRUE, by.x='dest', by.y='faa', suffixes=c('','.dest') )
data.joined <- YX
```



## Add a categorical variable for  arr_delay >= 22 minutes. It is called arrival_delayed  

```{r}
data.joined$arrival_delayed <- ifelse(data.joined$arr_delay >= 22, 1,0)
```


## Remove entries with NA values for Independant variables that we want to use in the model

```{r}

data.joined <- data.joined  %>% 
  filter(!is.na(dep_delay))   %>% 
  filter(!is.na(dest))   %>% 
  filter(!is.na(origin))   %>% 
  filter(!is.na(year))   %>% 
  filter(!is.na(month))   %>% 
  filter(!is.na(day))   %>% 
  filter(!is.na(hour))   %>% 
  filter(!is.na(tailnum))   %>% 
  filter(!is.na(sched_dep_time))   %>% 
  filter(!is.na(sched_arr_time))   %>% 
  filter(!is.na(flight))   %>% 
  filter(!is.na(distance))   %>% 
  filter(!is.na(year.pl))   %>% 
  filter(!is.na(minute))   %>% 
  filter(!is.na(year.pl))   %>% 
  filter(!is.na(type))   %>% 
  filter(!is.na(manufacturer))   %>% 
  filter(!is.na(model))   %>% 
  filter(!is.na(engines))   %>% 
  filter(!is.na(seats))   %>% 
  filter(!is.na(engine))   %>% 
  filter(!is.na(temp))   %>% 
  filter(!is.na(dewp))   %>% 
  filter(!is.na(humid))   %>% 
  filter(!is.na(wind_dir))   %>% 
  filter(!is.na(wind_speed))   %>% 
  filter(!is.na(wind_gust))   %>% 
  filter(!is.na(precip))   %>% 
  filter(!is.na(pressure))   %>% 
  filter(!is.na(visib))   %>% 
  filter(!is.na(name))   %>% 
  filter(!is.na(lat))   %>% 
  filter(!is.na(lon))   %>% 
  filter(!is.na(tz))   %>%
  filter(!is.na(name.dest))   %>% 
  filter(!is.na(lat.dest))   %>% 
  filter(!is.na(lon.dest))   %>% 
  filter(!is.na(alt.dest))   %>%
  filter(!is.na(tz.dest)) 
```

## Create Training and Test Samples


```{r}
data.joined.training <- sample_frac(data.joined, .75)
data.joined.testing <- sample_frac(data.joined, .5)
```

## Generate a Logistic Model

```{r}

logit.fit <- glm(arrival_delayed ~  dep_delay + dest + origin + year + month + day + hour + sched_dep_time + sched_arr_time + carrier + distance + year.pl + type  + engines + seats + engine + temp + dewp + humid + wind_dir + wind_speed + wind_gust +precip + pressure + visib + lat + lon + alt + tz  + lat.dest + lon.dest + alt.dest + tz.dest, data = data.joined.training, family = binomial)
#summary(logit.fit)

```

## Predict

```{r}
prob = plogis(predict(logit.fit, data.joined.testing))
```


# Question:

Is this a good model?  (Write your answer here.)

To answer this let's first calculate the misclassifications, Specificity and Sensitivity of the model. 

## Model Evaluation 

### Decide on optimal prediction probability cutoff for the model

Using the InformationValue::optimalCutoff  we determine the optimal cut-off
```{r}
optimalCutoff(data.joined.testing$arrival_delayed, prob)[1] 
optCutOff <- optimalCutoff(data.joined.testing$arrival_delayed, prob)[1] 
```

### Mis-classification rate

```{r}
misClassError(data.joined.testing$arrival_delayed, prob, threshold = optCutOff)
```

### Sensitivity

```{r}
sensitivity(data.joined.testing$arrival_delayed, prob, threshold = optCutOff)
```

### Specificity
```{r}
specificity(data.joined.testing$arrival_delayed, prob, threshold = optCutOff)
```

### Confusion Matrix
```{r}
confusionMatrix(data.joined.testing$arrival_delayed, prob, threshold = optCutOff)
```

## Conclusion
Based on the Confusion Matrix results, high Specificity and high Sensitivity and very low mis-classification errors this is a fairly good model, with high accuracy, sensitivity and specivity.


# PART B:

Your model should be good at explaining tardiness. Now, assume that your job is to predict arrival delays a month in advance. You can no longer use all the features in your model. Retrain your model using only features that will be *known* only a month in advance of the departure time.  Show all steps as above.


```{r}
logit.fit <- glm(arrival_delayed ~  dest + origin + sched_dep_time + sched_arr_time + carrier + distance + lat + lon + alt + tz  + lat.dest + lon.dest + alt.dest + tz.dest, data = data.joined.training, family = binomial(link="logit"))
```


## Model Evaluation 
Let's see how good this model is using a Confusion Matrix.


### Decide on optimal prediction probability cutoff for the model

Using the InformationValue::optimalCutoff  we determine the optimal cut-off
```{r}
prob = plogis(predict(logit.fit, data.joined.testing))
optimalCutoff(data.joined.testing$arrival_delayed, prob)[1] 
optCutOff <- optimalCutoff(data.joined.testing$arrival_delayed, prob)[1] 
```

### Mis-classification rate

```{r}
misClassError(data.joined.testing$arrival_delayed, prob, threshold = optCutOff)
```

### Sensitivity
Note that the **Sensitivity** is very low
```{r}
sensitivity(data.joined.testing$arrival_delayed, prob, threshold = optCutOff)
```

### Specificity
```{r}
specificity(data.joined.testing$arrival_delayed, prob, threshold = optCutOff)
```

### Confusion Matrix
```{r}
confusionMatrix(data.joined.testing$arrival_delayed, prob, threshold = optCutOff)
```

## Conclusion
Based on the Confusion Matrix results, and a very low Sensitivity, we can conclude that this is **NOT** a Model Fit.


```{r}
#prob = predict(logit.fit, data.joined.sample, type="response")
#prob = plogis(predict(logit.fit, data.joined.testing))
#prob<-ifelse(prob> 0.5,1,0)
#data.joined.testing$prob = prob

```

