---
title: "Improving Model Perfromance / Tuning Parameters"
author: "Saqib Ali"
date: '`r Sys.Date()`'
output:
  pdf_document: default
  html_document: default
---

```{r init, warning=FALSE, echo=FALSE, message=FALSE}
library(rpart)
library(caret)
library(data.table)
library(magrittr)
library(dplyr)


# .. Additional libraries
```


## Load the data

```{r}
data.joined <- readRDS(file="/home/saqib/ml_at_berkeley/CSX460/04-logistic-regression/04-exercise-nycflights-logistic/flightdata.Rda")


```


## Add a categorical variable for  arr_delay >= 22 minutes. It is called arrival_delayed  

```{r}
data.joined$arrival_delayed <- factor(ifelse(data.joined$arr_delay >= 22, 1,0))



```



## Filter out rows with NAs for arrival_delayed

```{r}

data.joined <- data.joined  %>% filter(!is.na(arrival_delayed))   
data.joined.training <- sample_frac(data.joined, .1)
data.joined.testing <- sample_frac(data.joined, .1)

#data.joined <- data.joined[, speed:=NULL]
```


## Tuning Parameter

Generically and regardless of model type, what are the purposes of a model
tuning parameters?

```
```

## Caret Models

This assignment demonstrates the use of caret for constructing models. Each
model should be built and compared using using `Kappa` as the performance
metric calculated using 10-fold repeated cross-validation with 3 folds.

Using the rectangular data that you created for the NYCFlights to create a model
for arr_delay >= 15 minutes.

- glm
- rpart
- knn
- C50
- randomForest
- adaBoost
- Two methods of your choice from the Caret Model List (you will need to install any dependencies)

Save the caret objects with the names provided.

```{r}

# Your work here

fitControl <- trainControl(method = "cv", number = 2)

fit.glm <- train(arrival_delayed ~ dep_delay + dest + origin + year + month + day + hour + sched_dep_time + sched_arr_time + carrier + distance + year.pl + type  + engines + seats + engine + temp + dewp + humid + wind_dir + wind_speed + wind_gust +precip + pressure + visib + lat + lon + alt + tz  + lat.dest + lon.dest + alt.dest + tz.dest, data=data.joined.training, method = "glm", trControl = fitControl)

fit.knn <- train(arrival_delayed ~ dep_delay + dest + origin + year + month + day + hour + sched_dep_time + sched_arr_time + carrier + distance + year.pl + type  + engines + seats + engine + temp + dewp + humid + wind_dir + wind_speed + wind_gust +precip + pressure + visib + lat + lon + alt + tz  + lat.dest + lon.dest + alt.dest + tz.dest, data=data.joined, method = "knn", trControl = fitControl)

fit.rpart <- train(arrival_delayed ~ dep_delay + dest + origin + year + month + day + hour + sched_dep_time + sched_arr_time + carrier + distance + year.pl + type  + engines + seats + engine + temp + dewp + humid + wind_dir + wind_speed + wind_gust +precip + pressure + visib + lat + lon + alt + tz  + lat.dest + lon.dest + alt.dest + tz.dest, data=data.joined.training, method = "rpart", trControl = fitControl)

fit.rf <- train(arrival_delayed ~ dep_delay + dest + origin + year + month + day + hour + sched_dep_time + sched_arr_time + carrier + distance + year.pl + type  + engines + seats + engine + temp + dewp + humid + wind_dir + wind_speed + wind_gust +precip + pressure + visib + lat + lon + alt + tz  + lat.dest + lon.dest + alt.dest + tz.dest, data=data.joined.training, method = "rf", trControl = fitControl)

```



## GLM
```{r}
pred <- as.vector(ifelse(predict(fit.glm, newdata=data.joined, type="prob")[,"1"]<.5, 0, 1))
probsTest <- predict(fit.glm, data.joined.testing, type = "prob")
pred      <- factor( ifelse(probsTest[, "1"] > 0.5, "1", "0") )
confusionMatrix(pred, data.joined.testing$arrival_delayed)

```


## KNN

```{r}
pred <- as.vector(ifelse(predict(fit.knn, newdata=data.joined, type="prob")[,"1"]<.5, 0, 1))
probsTest <- predict(fit.knn, data.joined.testing, type = "prob")
pred      <- factor( ifelse(probsTest[, "1"] > 0.5, "1", "0") )
confusionMatrix(pred, data.joined.testing$arrival_delayed)

```


## rpart


```{r}
pred <- as.vector(ifelse(predict(fit.rpart, newdata=data.joined, type="prob")[,"1"]<.5, 0, 1))
probsTest <- predict(fit.rpart, data.joined.testing, type = "prob")
pred      <- factor( ifelse(probsTest[, "1"] > 0.5, "1", "0") )
confusionMatrix(pred, data.joined.testing$arrival_delayed)

```

## Random Forest 

```{r}
pred <- as.vector(ifelse(predict(fit.rf, newdata=data.joined, type="prob")[,"1"]<.5, 0, 1))
probsTest <- predict(fit.rf, data.joined.testing, type = "prob")
pred      <- factor( ifelse(probsTest[, "1"] > 0.5, "1", "0") )
confusionMatrix(pred, data.joined.testing$arrival_delayed)

```


Compare the  models?

Which is best?  Why?

```
```
