---
title: "Classification Metrics"
author: "Saqib Ali"
date: "May 09, 2017"
output:
  pdf_document: default
  html_document: default
---


**KNIT YOUR DOCUMENT AS *HTML* AND SUBMIT IT AND THE `Rmd` file.** 
 

## Classification Metrics Functions 

Write functions of `y` and `y_hat` to calculate the following classification metrics

* True Positive Rate  
* False Positive Rate
* True Negative Rate  
* False Negative Rate 
* Sensitivity 
* Specificity 
* Recall 
* Precision
* Prevalence 
* Accuracy
* Kappa

```{r "definitions", echo=FALSE}
library(magrittr)
library("tidyverse")
library("ggplot2")
library("lubridate")
library("dplyr")
library("data.table")

library(readr)
library(ggplot2)
library(magrittr)
library(lubridate)
library(stringr)
library(dplyr)
#library(caret)
library(InformationValue)


library(data.table)

data.joined <- readRDS(file="/home/saqib/ml_at_berkeley/CSX460/04-logistic-regression/04-exercise-nycflights-logistic/flightdata.Rda")
data.joined$arrival_delayed <- ifelse(data.joined$arr_delay >= 22, 1,0)

data.joined <- data.joined  %>% filter(!is.na(arrival_delayed))   

data.joined.training <- sample_frac(data.joined, .75)
data.joined.testing <- sample_frac(data.joined, .5)
logit.fit <- glm(arrival_delayed ~  dest + origin + sched_dep_time + sched_arr_time + carrier + distance + lat + lon + alt + tz  + lat.dest + lon.dest + alt.dest + tz.dest, data = data.joined.training, family = binomial(link="logit"))
logit.fit <- glm(arrival_delayed ~  dep_delay + dest + origin + year + month + day + hour + sched_dep_time + sched_arr_time + carrier + distance + year.pl + type  + engines + seats + engine + temp + dewp + humid + wind_dir + wind_speed + wind_gust +precip + pressure + visib + lat + lon + alt + tz  + lat.dest + lon.dest + alt.dest + tz.dest, data = data.joined.training, family = binomial)



```

```{r}
prob = plogis(predict(logit.fit, data.joined.testing))
prob<-ifelse(prob> 0.5,1,0)
data.joined.testing$prob = prob


confusionMatrix(data.joined.testing$arrival_delayed, prob)
y = data.joined.testing$arrival_delayed
y_hat = data.joined.testing$prob
```
```{r}



#TP <- sum((y==0)&(y==y_hat))
#TN <- sum((y==1)&(y==y_hat))
#FP <- sum((y_hat==0)&(y==1))
#FN <- sum((y_hat==1)&(y==0))

Accuracy <- function(y, y_hat){
  (sum((y==0)&(y==y_hat)) + sum((y==1)&(y==y_hat))) / (sum((y==0)&(y==y_hat)) + sum((y==1)&(y==y_hat)) + sum((y_hat==0)&(y==1)) + sum((y_hat==1)&(y==0)))
}

TruePositiveRate <- function(y, yhat){
 sum((y==0)&(y==y_hat))/(sum((y==0)&(y==y_hat)) + sum((y_hat==1)&(y==0)))
  
}  


FalsePositiveRate <- function(y, y_hat){
  sum((y_hat==0)&(y==1))/(sum((y_hat==0)&(y==1)) + sum((y==1)&(y==y_hat)))
}


TrueNegativeRate  <- function(y, y_hat){
  sum((y==1)&(y==y_hat))/(sum((y==1)&(y==y_hat)) + sum((y_hat==0)&(y==1)))

}


FalseNegativeRate  <- function(y, y_hat){
 sum((y_hat==1)&(y==0))/(sum((y_hat==1)&(y==0)) + sum((y==0)&(y==y_hat)))
}

Sensitivity <- TruePositiveRate
Specificity <- TrueNegativeRate

Recall <- TruePositiveRate


Precision <- function(y, y_hat){
  sum((y==0)&(y==y_hat)) / (sum((y==0)&(y==y_hat))+sum((y_hat==0)&(y==1)))
}

Prevalence <- function(y, y_hat){
  sum(y==1)/length(y)
  
}

# Kappa


```


```{r}

Accuracy(y,y_hat)
1-misClassError(data.joined.testing$arrival_delayed, prob)

TrueNegativeRate(y,y_hat)
sensitivity(data.joined.testing$arrival_delayed, prob)

TruePositiveRate(y,y_hat)
FalsePositiveRate(y,y_hat)
FalseNegativeRate(y,y_hat)
Sensitivity(y,y_hat)
Specificity(y,y_hat)
specificity(data.joined.testing$arrival_delayed, prob)

Recall(y,y_hat)

Precision(y,y_hat)

Prevalence(y,y_hat)

```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


## Usage 

Use the function(s) above to calculate the metrics for your late arrival model that
you created last week. 

```{r "evaluations"}


```

