bv---
title: "Assingment NYC Flight Data"
author: "Your Name Here"
date: "October 5, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library("tidyverse")
library("ggplot2")
library("lubridate")
library("dplyr")
```

## NYC Flight Data

Your job is to create a "rectangular" table useful for modeling from flight data. 
In the folder `02-fundamentals/data` there are four CSV files: `flights.csv`, 
`airports.csv`, `planes.csv` and `weather.csv`.  Join/merge these tables such 
that there is one rectangular table with one row for each flight. 


Put code in each of the sections provided.

## 1. Read Data

Using the `readr` package read the `flights` data.

```{r}

flights <- read.csv("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/flights.csv")


str(flights)
planes <- read.csv("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/planes.csv")
airports <- read.csv("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/airports.csv")
weather <- read.csv("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/weather.csv")


```

## Numeric Variables  

Plot a histogram of arrival delays and departure delays 

```{r}

library(reshape2)



delays = flights[,c("dep_delay", "arr_delay")]

ggplot(data=flights, aes(flights$dep_delay)) + geom_histogram(binwidth = .5)
ggplot(data=flights, aes(flights$arr_delay)) + geom_histogram(binwidth = .5)


#ggplot(data = iris2) +
#  geom_histogram(aes(x = value, y=(..count..)/sum(..count..), fill=variable), 
#                 alpha=0.3, binwidth=2, position="identity")


```

## Scatterplot

Plot a scatterplot of date vs arrival delay 

```{r}
form.dates <- ymd(paste(flights$year, flights$month, flights$day, sep = "-"))
flights$date <- form.dates

ggplot(flights, aes(x=date, y=arr_delay)) +
    geom_point()  

```



## Create tables for Categorical Variables

Create a `table`, counting the flights by origin airports and carrier.

```{r}

flights %>% group_by(origin, carrier) %>% summarise(total.count=n())


```


## Join Data

**Read** in the other data sets. Use methods to join the data together to create a single table containing one record per row.  (HINT: see `?data.table` or `?dplyr::join`)

```{r}

data.joined <- flights %>% left_join(planes, by = "tailnum") %>%  left_join(airports, c("dest"="faa")) %>%  left_join(airports, c("origin"="faa")) %>% left_join(weather, by = c('time_hour'='time_hour', 'origin'='origin'))

str(data.joined)


```



