---
title: "Assingment NYC Flight Data"
author: "Your Name Here"
date: "October 5, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library("tidyverse")
library("ggplot2")
library("lubridate")
library("dplyr")
```

## NYC Flight Data

Your job is to create a "rectangular" table useful for modeling from flight data. 
In the folder `02-fundamentals/data` there are four CSV files: `flights.csv`, 
`airports.csv`, `planes.csv` and `weather.csv`.  Join/merge these tables such 
that there is one rectangular table with one row for each flight. 


Put code in each of the sections provided.

## 1. Read Data

Using the `readr` package read the `flights` data.

```{r}

flights <- read.csv("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/flights.csv")

planes <- read.csv("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/planes.csv")
airports <- read.csv("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/airports.csv")
weather <- read.csv("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/weather.csv")


#flights <- flights[1:10,]
print(flights)
```

## Numeric Variables  

Plot a histogram of arrival delays and departure delays 

```{r}

ggplot(flights, aes(x = dep_delay, x=arr_delay)) +
        geom_histogram()
```

## Scatterplot

Plot a scatterplot of date vs arrival delay 

```{r}

ggplot(flights, aes(x=time_hour, y=arr_delay)) +
    geom_point(shape=1)  

```



## Create tables for Categorical Variables

Create a `table`, counting the flights by origin airports and carrier.

```{r}

flights %>% group_by(origin, carrier) %>% summarise(total.count=n())


```


## Join Data

**Read** in the other data sets. Use methods to join the data together to create a single table containing one record per row.  (HINT: see `?data.table` or `?dplyr::join`)

```{r}


#flights %>% left_join(planes, by = "tailnum") %>% left_join(weather, by ="time_hour") %>%  left_join(airports, c("dest"="faa"))

df_flights <- read.csv("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/flights.csv")

df_planes <- read.csv("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/planes.csv")
df_airports <- read.csv("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/airports.csv")
df_weather <- read.csv("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/weather.csv")


```



