---
title: "Recursive Partitioning / Decision Trees using CARET"
author: "Saqib Ali"
date: "May 16th, 2017"
output:
  html_document: default
  pdf_document: default
---

```{r init, warning=FALSE, echo=FALSE, message=FALSE}
library(rpart)
library(caret)
library(data.table)
library(magrittr)
library(dplyr)


# .. Additional libraries
```


## Load the data

```{r}
data.joined <- readRDS(file="/home/saqib/ml_at_berkeley/CSX460/04-logistic-regression/04-exercise-nycflights-logistic/flightdata.Rda")


```


## Add a categorical variable for  arr_delay >= 22 minutes. It is called arrival_delayed  

```{r}
data.joined$arrival_delayed <- factor(ifelse(data.joined$arr_delay >= 22, 1,0))



```



## Filter out rows with NAs for arrival_delayed

```{r}

data.joined <- data.joined  %>% filter(!is.na(arrival_delayed))   

#data.joined <- data.joined[, speed:=NULL]
```


## Split the Dataset in Training and Test datasets
```{r}

data.joined.training <- sample_frac(data.joined, .75)
data.joined.testing <- sample_frac(data.joined, .5)

```


## Exercise 1: caret/logistic regression (5 points)

Rebuild your logistic regression model from the previous week, this time using the `caret` package. 

- Calculate the training or apparent performance of the model. 
- Calculate an unbiased measure of performance 
- Create a ROC Curve for your model

Show all work.


### Train the Linear Regression Model using CARET
```{r logistic model}
# Your Work Here

#data.joined <- data.joined[, speed:=NULL]


#lapply(data.joined, levels)
#(l <- sapply(data.joined, function(x) is.factor(x)))

fitControl <- trainControl(method = "cv", number = 2)

glmFit <- train(arrival_delayed ~ dep_delay + dest + origin + year + month + day + hour + sched_dep_time + sched_arr_time + carrier + distance + year.pl + type  + engines + seats + engine + temp + dewp + humid + wind_dir + wind_speed + wind_gust +precip + pressure + visib + lat + lon + alt + tz  + lat.dest + lon.dest + alt.dest + tz.dest, data=data.joined, method = "glm", trControl = fitControl)





```

### Model Performance
```{r}

pred <- as.vector(ifelse(predict(glmFit, newdata=data.joined, type="prob")[,"1"]<.5, 0, 1))
probsTest <- predict(glmFit, data.joined.testing, type = "prob")
pred      <- factor( ifelse(probsTest[, "1"] > 0.5, "1", "0") )
confusionMatrix(pred, data.joined.testing$arrival_delayed)


```

### Plot the ROC Curve
```{r}
library(pROC)
probsTrain <- predict(glmFit, data.joined.training, type = "prob")
rocCurve   <- roc(response = data.joined.training$arrival_delayed,
                      predictor = probsTrain[, "1"],
                      levels = rev(levels(data.joined.training$arrival_delayed)))
plot(rocCurve, print.thres = "best")

```

## Exercise 2: caret/rpart (5 points)

Using the `caret` and `rpart` packages, create a **classification** model for flight delays using your NYC FLight data. Your solution should include:

- The use of `caret` and `rpart` to train a model.
- An articulation of the the problem your are 
- An naive model
- An unbiased calculation of the performance metric
- A plot of your model -- (the actual tree; there are several ways to do this)
- A discussion of your model 



Show and describe all work

```{r rpart model}
# Your Work Here

fitControl <- trainControl(method = "cv", number = 2)

rpartFit <- train(arrival_delayed ~ dep_delay + dest + origin + year + month + day + hour + sched_dep_time + sched_arr_time + carrier + distance + year.pl + type  + engines + seats + engine + temp + dewp + humid + wind_dir + wind_speed + wind_gust +precip + pressure + visib + lat + lon + alt + tz  + lat.dest + lon.dest + alt.dest + tz.dest, data=data.joined, method = "rpart", trControl = fitControl)




```

### Model Performance

```{r}
pred <- as.vector(ifelse(predict(rpartFit, newdata=data.joined, type="prob")[,"1"]<.5, 0, 1))
probsTest <- predict(rpartFit, data.joined.testing, type = "prob")
pred      <- factor( ifelse(probsTest[, "1"] > 0.5, "1", "0") )
confusionMatrix(pred, data.joined.testing$arrival_delayed)
```


### Decision Tree Plot
```{r}
library(rpart.plot)


rpart.plot(rpartFit$finalModel)
```


### Questions:

- Discuss the difference between the models and why you would use one model over the other?


Logistic Regression performed better than Decision Trees.