---
title: "NYCFlights: Arrival Delay Regression Model (Open Ended Variation)"
author: "[Your Name]"
date: "[Today's Date]"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library("tidyverse")
library("ggplot2")
library("lubridate")
library("dplyr")
library("data.table")

library(readr)
library(ggplot2)
library(magrittr)
library(lubridate)
library(stringr)
library(dplyr)
library(data.table)

```

# NYCFlights Model

Using the rectangular data that you created last week create a model for `arr_delay`. Follow a well defined process for creating a model as discussed in class or from one of the readings. 

Break the process down into steps and create a codeblock for each step. 
 - Use as many steps you feel as necessary
 - Concisely explain each step 
 - Show all work
 
(You can start with the data that you created last week.)


## Step 1: Read the data

```{r "Step 1: [Step Name]" }





flightsDT  <- fread("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/flights.csv") 
planesDT   <- fread("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/planes.csv")  %>% setkey(tailnum)
airportsDT <- fread("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/airports.csv")  %>% setkey(faa) 
weatherDT  <- fread("~/ml_at_berkeley/CSX460/02-building-blocks/02-exercise-nycflights/data/weather.csv") %>%   setkey(origin,year,month,day,hour)



```
## Step 2: Join the datasets and analyze the structure
```{r}



#data.joined <- flights %>% left_join(planes, by = c('tailnum' = 'tailnum')) %>%  left_join(airports, c("dest"="faa"))  %>% left_join(weather, by = c('year'='year', 'month'='month', 'day'='day', 'dep_hour'='hour', 'origin'='origin')) 


YX <- flightsDT 
YX %<>% merge( planesDT, all.x = TRUE, by='tailnum', suffixes=c('','.pl') )
YX %<>% merge( weatherDT, all.x = TRUE, by=c('origin','year','month','day','hour'), suffixes=c('','.we') )
YX %<>% merge( airportsDT, all.x = TRUE, by.x='origin', by.y='faa', suffixes=c('','.orig') )
YX %<>% merge( airportsDT, all.x = TRUE, by.x='dest', by.y='faa', suffixes=c('','.dest') )


data.joined <- YX
dim(data.joined)

str(data.joined)
```

## Step 3: Get sample training data (50% of the entire dataset)
```{r}
data.joined.sample <- sample_frac(data.joined, .5)



```


## Step 4:

Remove entries with NA values for Independant variables that we want to use in the model



```{r}

dim(data.joined.sample)
data.joined.sample <- data.joined.sample  %>% 
  filter(!is.na(dep_delay))   %>% 
  filter(!is.na(dest))   %>% 
  filter(!is.na(origin))   %>% 
  filter(!is.na(year))   %>% 
  filter(!is.na(month))   %>% 
  filter(!is.na(day))   %>% 
  filter(!is.na(hour))   %>% 
  filter(!is.na(tailnum))   %>% 
  filter(!is.na(sched_dep_time))   %>% 
  filter(!is.na(sched_arr_time))   %>% 
  filter(!is.na(flight))   %>% 
  filter(!is.na(distance))   %>% 
  filter(!is.na(year.pl))   %>% 
  filter(!is.na(minute))   %>% 
  filter(!is.na(year.pl))   %>% 
  filter(!is.na(type))   %>% 
  filter(!is.na(manufacturer))   %>% 
  filter(!is.na(model))   %>% 
  filter(!is.na(engines))   %>% 
  filter(!is.na(seats))   %>% 
  filter(!is.na(engine))   %>% 
  filter(!is.na(temp))   %>% 
  filter(!is.na(dewp))   %>% 
  filter(!is.na(humid))   %>% 
  filter(!is.na(wind_dir))   %>% 
  filter(!is.na(wind_speed))   %>% 
  filter(!is.na(wind_gust))   %>% 
  filter(!is.na(precip))   %>% 
  filter(!is.na(pressure))   %>% 
  filter(!is.na(visib))   %>% 
  filter(!is.na(name))   %>% 
  filter(!is.na(lat))   %>% 
  filter(!is.na(lon))   %>% 
  filter(!is.na(tz))   %>%
  filter(!is.na(name.dest))   %>% 
  filter(!is.na(lat.dest))   %>% 
  filter(!is.na(lon.dest))   %>% 
  filter(!is.na(alt.dest))   %>%
  filter(!is.na(tz.dest)) 

  
  
  
  

  
  
  
  dim(data.joined.sample)




```

## Step 5: Create a model with the relevant Features (independant variables). Note: We can not use dep_delay as it will be not available to us in the real world for predicting arrival delays

```{r}

fit <- lm(arr_delay ~  dest + origin + year + month + day + hour + tailnum + sched_dep_time + sched_arr_time + carrier + flight + distance + year.pl + type + manufacturer + model + engines + seats + engine + temp + dewp + humid + wind_dir + wind_speed + wind_gust +precip + pressure + visib + name + lat + lon + alt + tz + name.dest + lat.dest + lon.dest + alt.dest + tz.dest, data=data.joined.sample)

#scatter.smooth(x=data.joined.sample$year, y=data.joined.sample$dep_delay)
```




Step 6: Create a histogram of the Residuals and Summary of the Model. 
```{r}
hist(resid(fit))
summary(fit)
```


# Question: Is this a good model?  (Write your answer here.)
This is a not a good model. Based on the extremely low R-squared, we can say that is not a good model.








```{r}
fit <- lm(arr_delay ~  dest + origin + year + month + day + hour + sched_dep_time + sched_arr_time + carrier + distance + year.pl + type + manufacturer + engines + seats + engine + temp + dewp + humid + wind_dir + wind_speed + wind_gust +precip + pressure + visib + name + lat + lon + alt + tz + name.dest + lat.dest + lon.dest + alt.dest + tz.dest, data=data.joined.sample)



```

